---
title: "nottingham-temperatures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nottingham-temperatures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(reproseries)
library(magrittr)
```

The `reproseries` package itself is not sufficient, we need `magrittr` to pipe functions together.

# Introduction

This vignette shows what you can do with the functions in the `reproseries` package. Code is not hidden so you can follow along nicely.

# Data

The functions are written to be used on the built-in `nottem` dataset. This dataset looks like this:

```{r}
temperatures <- nottem %>%
  temperatureSeriesToDf

temperatures %>% str
```

We first transform `nottem` from a time series in a dataframe object. This way `str` gives us a clearer overview of what the dataset looks like.

# Analysis phases

## Tidy

The raw `nottem` dataset is not tidy since variables are spread across columns. There are 3 variables (month, year and temperature) but there are 12 columns (one for each month). Let's fix this:

```{r}
temperatures <- temperatures %>% 
  tidyTemperatures

temperatures %>% str
```

Now we have the 3 variables we need. Month and year are clear, value refers to the temperature in fahrenheit. In the next step we'll deal with this vague `value` variable.

## Transform

We're no Americans so transform fahrenheits to celsius.

```{r}
temperatures <- temperatures %>% 
  addCelsiusColumn

temperatures %>% str
```

In preparation of the modeling and visualisation we already calculate an average temperature per year.

```{r}
avg_temperature_by_year <- temperatures %>% 
  summariseAvgTemperatureByYear

avg_temperature_by_year %>% str
```

## Visualize

How did the average temperature evolve over the years?

```{r}
avg_temperature_by_year %>%
  plotAvgTemperaturesByYear
```

## Model

Is there a trend in the average temperatures?

```{r}
avg_temperature_by_year %>%
  calculateAutoCorrelation
```

0.10 is rather low so there doesn't seem to be any trend.

# Summary

All the functions available in the `reproseries` package have been introduced above.
